---
title: "rtabulator"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

rtabulator provides R bindings for [Tabulator JS](https://tabulator.info/).

The goal of rtabulator is to make it easy to create elegant and interactive tables in markdown documents and [Shiny](https://shiny.posit.co/) applications. Furthermore, it supports creating spreadsheets with multiple sheets.

```{r setup}
library(rtabulator)
```

## Basic Usage

To render a table just pass a data frame to `tabulator()`:

```{r}
tabulator(airquality)
```

In this case, the column definitions are automatically created for you. The horizontal alignment for character columns is set to _left_  while numeric columns are aligned _right_.

It is also possible to pass a data URL:

```{r}
data_url <- "https://raw.githubusercontent.com/eodaGmbH/rtabulator/main/data-raw/titanic.csv"

tabulator(
  data_url,
  col_select = c("PassengerId", "Name", "Pclass", "Fare", "Survived")
)
```

## Setup Options

With `tabulator_options()` you can customize your table (or spreadsheet):

```{r}
setup <- tabulator_options(
  row_height = 50,
  selectable_rows = FALSE,
  edit_trigger_event = "click"
)

tabulator(iris, setup, editable = TRUE)
```

## Pagination

```{r}
setup <- tabulator_options(
  pagination = TRUE,
  pagination_size = 10,
  pagination_size_selector = c(10, 20, 50)
)

tabulator(USArrests, setup)
```

***

**Note**

If you prefer the pipe style, you can also use the helper function `set_options_pagination()`.

***

## Column Formatters

To customize your columns, the easiest way is to use the `set_formatter_*()` functions: 

```{r}
tabulator(airquality) |>
  set_formatter_progress("Temp", legend = TRUE, legend_align = "left") |>
  set_formatter_traffic_light("Ozone") |>
  set_tooltip("Ozone")
```

With `for_each_col()` you can apply a formatter and any other function that updates the column settings to multiple columns:

```{r}
numeric_cols <- c("Sepal_Length", "Sepal_Width", "Petal_Length", "Petal_Width")

tabulator(iris) |>
  for_each_col(
    numeric_cols,
    set_formatter_progress,
    legend = TRUE,
    legend_align = "left",
    min = 0,
    max = 10,
    color = c("yellow", "orange", "green")
  ) |>
  for_each_col(numeric_cols, set_tooltip)
```

## Multi Column headers

You can create multi column headers with `set_multi_column_header()`:

```{r}
headers <- list(
  Sepal = c("Sepal_Width", "Sepal_Length"),
  Petal = c("Petal_Width", "Petal_Length")
)

tabulator(iris) |>
  set_header_filter("Species", "list", placeholder = "Select Species") |>
  set_multi_column_header(headers)
```

***

**Note**

If further functions update the column definitions, `set_multi_column_header()` must be the last one. Therefore, in the example above `set_header_filter()` is called before we set the multi column header.

***

## Groupings

```{r}
setup <- tabulator_options(
  group_by = c("Pclass"),
  group_start_open = FALSE,
  group_toggle_element = "header"
)

titanic_df <- titanic(c("Pclass", "Sex", "Age", "Survived"))

tabulator(titanic_df, setup) |>
  set_formatter_traffic_light("Survived", color = c("red", "green"))
```

Multi level grouping is done by passing a vector as `group_by` parameter:

```{r}
setup <- tabulator_options(
  group_by = c("Sex", "Pclass"),
  group_start_open = c(TRUE, FALSE)
)

tabulator(titanic_df, setup)
```

***

**Note**

You can also use `set_options_group_by()` to set the group by options. 

***

## Calculations

With `set_calculation()` you can add calculations to the top or bottom of columns:

```{r}
tabulator(iris[4:5]) |>
  set_calculation("Petal_Width", "avg", pos = "top") |>
  set_calculation("Petal_Width", "max", precision = 1, pos = "bottom")
```

## Editors

Set `editable = TRUE` to make all columns of the table editable. In this case, the editor for numeric columns is set to `number` and all others to `input`.

```{r}
setup <- tabulator_options(edit_trigger_event = "click")

tabulator(iris, setup, editable = TRUE)
```

To customize editor settings and validate user input use `set_editor()`:

```{r}
tabulator(iris[, c(2, 5)], setup) |>
  set_editor(
    column = "Species",
    editor = "list",
    values_lookup = "active",
    clearable = TRUE,
    autocomplete = TRUE
  ) |>
  set_editor(
    column = "Sepal_Width",
    editor = "number",
    min = 0,
    max = 10,
    step = 0.1,
    validator = c("min:0", "max:10")
  )
```

## Spreadsheets

To create an empty spreadsheet just pass an empty list as `data` parameter and `spreadsheet = TRUE` to `tabulator()`:

```{r}
tabulator(data = list(), spreadsheet = TRUE)
```

The data format for spreadsheets is a list of lists or vectors where each entry represents a row in the spreadsheet:

```{r}
spreadsheet_data <- list(
  c(1, 2, 3),
  c(2, 3, 4)
)
```

To use multiple sheets pass a list of sheets created with`spreadsheet_def()`. In this case, data is passed with the `data` parameter of the sheet definitions:

```{r}
sheets <- list(
  spreadsheet_def(title = "Lee", data = spreadsheet_data),
  spreadsheet_def(title = "Morgan", data = list()) # Empty spreadsheet
)
setup <- tabulator_options(
  spreadsheet = TRUE,
  spreadsheet_sheets = sheets,
  spreadsheet_sheet_tabs = TRUE,
  edit_trigger_event = "click"
)

tabulator(NULL, setup)
```
## Next Steps

* [API Reference](../reference/index.html)
* [Shiny Integration](../articles/shiny.html)
* [Themes](../articles/themes.html)
