---
title: "rtabulator"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

rtabulator provides R bindings for [Tabulator JS](https://tabulator.info/).

The goal of rtabulator is to make it easy to create elegant and interactive tables in markdown documents and [Shiny](https://shiny.posit.co/) applications. Furthermore, it support creating spreadsheets with multiple sheets.

```{r setup}
library(rtabulator)
```

## Basic Usage

To render a table just pass a data frame to the `tabulator` function:

```{r}
tabulator(airquality)
```

In this case, the column definitions are automatically created for you. The horizontal alignment for
character columns is set to _left_  while numeric columns are aligned _right_.

## Setup Options

With `tabulator_options` you can customize your table (or spreadsheet):

```{r}
setup <- tabulator_options(
  pagination = TRUE,
  pagination_size = 10,
  pagination_size_selector = c(10, 20, 50)
)

tabulator(USArrests, setup)
```

## Column Formatters

To customize your columns the easiest way is to use the `set_formatter_*()` functions: 

```{r}
tabulator(airquality) |>
  set_formatter_progress("Temp", legend = TRUE, legend_align = "left") |>
  set_formatter_traffic_light("Ozone")
```

## Multi Column headers

You can create multi column headers with `set_multi_column_header`:

```{r}
headers <- list(
  Sepal = c("Sepal_Width", "Sepal_Length"),
  Petal = c("Petal_Width", "Petal_Length")
)

tabulator(iris) |>
  set_multi_column_header(headers)
```

## Spreadsheets

To create an empty spreadsheet just pass an empty list as `data` parameter and `spreadsheet = TRUE`
to the `tabulator` function:

```{r}
tabulator(data = list(), spreadsheet = TRUE)
```
